Dynamic Walking and Whole-Body Motion Planning for Humanoid Robots: an Integrated Approach

### 2012-06-20 paper.tex
constraint-solver -> \texttt{constraint-solver}
Minor rephrasing

\begin{algorithm}
\caption{constraint-solver($q$, f, $\varepsilon$): find $q$ such that $f(q) = 0$}
\label{algo:newton}
\begin{algorithmic}
\STATE $\alpha=.1$, $i=0$, $\alpha_{max} = .95$
\WHILE{$\|f(q)\| > \varepsilon$ and $i\leq 20$}
\STATE {\color{red} // $\left(.\right)^{+}$ denotes the Moore-Penrose pseudo-inverse}
\STATE $q \leftarrow$ $q - \alpha \left(\frac{\partial f}{\partial q}(q)\right)^{+} f(q)$
\STATE $i$ $\leftarrow$ $i+1$
\STATE {\color{red}// Make $\alpha$ tend toward $\alpha_{max}$}
\STATE $\alpha \rightarrow \alpha_{max} - .8*(\alpha_{max} - \alpha)$
\ENDWHILE
\IF {$\|f(q)\| \leq \varepsilon$}
\STATE return $q$
\ELSE
\STATE return failure
\ENDIF
\end{algorithmic}
\end{algorithm}

@@

The equivalent random  extension on a constrained manifold $\mathcal{M}$, 
defined by the constraint function $f$,
starts from a valid configuration $q_{near} \in \mathcal{M}$, and extends the tree
towards a random configuration $q_{rand}$, while keeping the constraints
satisfied. Extension attempts orthogonal to $\mathcal{M}$ are useless,
as newly added edges have to be included  in $\mathcal{M}$. To extend in directions
that follow the directions of $\mathcal{M}$, we rely on the framework of 
jacobian-based prioritized inverse kinematics. Algorithm~\ref{alg:constrained} presents 
the adaptation of the classic extend function, and Fig.~\ref{fig:gikrrt} illustrates this
extension. The idea is to first project $q_{rand}$ on the tangent
space to $\mathcal{M}$  at $q_{near}$. Let us call the projected configuration $q_{rand}'$.
$q_{rand}'$ is obtained by one step of prioritized inverse kinematics optimization. Let 
$q_{rand}''$  be the result of a call to $\text{constraint-solver}(q_{rand}',f,\epsilon)$.  It is 
the projection
of $q_{rand}'$ on $\mathcal{M}$. Instead of extending the tree from $q_{near}$ towards 
$q_{rand}$, the algorithm tries to extend from $q_{near}$ towards 
$q_{rand}''$\footnote{This presentation attempts to give a precise idea of the algorithm, without 
  focusing on technical details. Readers interested in the actual implementation can
  refer to \url{https://github.com/laas/hpp-constrained} and 
  \url{https://github.com/laas/hpp-constrained-planner}, where the corresponding open-source 
  code is available.}. While extending the tree,
the configurations along the new edge are automatically projected onto $\mathcal{M}$. These projections 
are not very costly if the edge is close to the constrained manifold.

@@

\begin{algorithm}[h]
  \caption{Constrained-Extend($\mathcal{T},q_{near},q_{rand},f,\epsilon$)}
  \label{alg:constrained}
  \begin{algorithmic}
    \STATE {\color{red} // $I$ denotes the identity matrix.}
    \STATE $q_{rand}' \leftarrow q_{near} +   
    \left(I -  
    \left(\frac{\partial f}{\partial q}(q_{near})\right)^{+}
    \frac{\partial f}{\partial q}(q_{near})\right) (q_{rand} - q_{near})$
    \IF {$\text{constraint-solver}(q_{rand}',f,\epsilon) = q_{rand}''$} 
    \STATE RRT::Extend($\mathcal{T},q_{near},q_{rand}''$)
    \ENDIF
  \end{algorithmic}
\end{algorithm}

@@

This result has been long known and used in motion planning, in particular in planning
for non-holonomic systems. A detailed proof of it can be found in
\cite{taix-94}. We will present a sketch of the proof to give an intuition about the 
corresponding algorithm.

@@

\subsubsection{Small-Time \textit{versus} Small-Space Controllability}
In the control theory literature, the property used is usually \textit{small-time controllability}, 
which states that for  all configurations  $q$, for  all times
$T>0$, the set of configurations accessible from $q$ in time less than
$T$ forms a  neighborhood of $q$. When accelerations and velocities are bounded,
small-time controllability implies small-space controllability, which is why 
a lot of motion planning previous work only refers to the sufficient small-time controllability
property. However, the reciprocate is not necessarily true:  a system can be 
small-space controllable and
not small-time, if the trajectories generated by its controller are arbitrarily long.
The important property, regarding
motion planning application, is small-space controllability, as 
Theorem~\ref{thm:ssc} shows. In the following, we show that legged robots
are small-space controllable, but  not that they are small-time controllable.
In fact, the  control method that we present does not follow the small-time controllability
property. For the sake of clarity, we have chosen to make the distinction between these
two controllability properties.

@@

This section  discusses walking robot small-space controllability. To clarify
the presentation, we consider a simplified model of a legged robot consisting of two feet
of null mass and a punctual mass free to move in three dimensions.
We do not consider 
the kinematic chains between the feet and the mass. The robot is walking on a flat terrain,
and the feet are assumed to have a positive surface. For our presentation, it is not 
necessary to consider the feet height, so the configuration space of the robot is:
\[
\mathcal{C} = SE(2) \times SE(2) \times \mathbb{R}^3
\]
It is of dimension 9.

>>>>

\begin{algorithm}
\caption{\texttt{constraint-solver}($q$, f, $\varepsilon$): find $q$ such that $f(q) = 0$}
\label{algo:newton}
\begin{algorithmic}
\STATE $\alpha=.1$, $i=0$, $\alpha_{max} = .95$
\WHILE{$\|f(q)\| > \varepsilon$ and $i\leq 20$}
\STATE {\color{red} // $\left(.\right)^{+}$ denotes the Moore-Penrose pseudo-inverse}
\STATE $q \leftarrow$ $q - \alpha \left(\frac{\partial f}{\partial q}(q)\right)^{+} f(q)$
\STATE $i$ $\leftarrow$ $i+1$
\STATE {\color{red}// Make $\alpha$ tend toward $\alpha_{max}$}
\STATE $\alpha \rightarrow \alpha_{max} - .8*(\alpha_{max} - \alpha)$
\ENDWHILE
\IF {$\|f(q)\| \leq \varepsilon$}
\STATE return $q$
\ELSE
\STATE return failure
\ENDIF
\end{algorithmic}
\end{algorithm}

@@

The equivalent random  extension on a constrained manifold $\mathcal{M}$, 
defined by the constraint function $f$,
starts from a valid configuration $q_{near} \in \mathcal{M}$, and extends the tree
towards a random configuration $q_{rand}$, while keeping the constraints
satisfied. Extension attempts orthogonal to $\mathcal{M}$ are useless,
as newly added edges have to be included  in $\mathcal{M}$. To extend in directions
that follow the directions of $\mathcal{M}$, we rely on the framework of 
jacobian-based prioritized inverse kinematics. Algorithm~\ref{alg:constrained} presents 
the adaptation of the classic extend function, and Fig.~\ref{fig:gikrrt} illustrates this
extension. The idea is to first project $q_{rand}$ on the tangent
space to $\mathcal{M}$  at $q_{near}$. Let us call the projected configuration $q_{rand}'$.
$q_{rand}'$ is obtained by one step of prioritized inverse kinematics optimization. Let 
$q_{rand}''$  be the result of a call to $\texttt{constraint-solver}(q_{rand}',f,\epsilon)$.  It is 
the projection
of $q_{rand}'$ on $\mathcal{M}$. Instead of extending the tree from $q_{near}$ towards 
$q_{rand}$, the algorithm tries to extend from $q_{near}$ towards 
$q_{rand}''$\footnote{This presentation attempts to give a precise idea of the algorithm, without 
  focusing on technical details. Readers interested in the actual implementation can
  refer to \url{https://github.com/laas/hpp-constrained} and 
  \url{https://github.com/laas/hpp-constrained-planner}, where the corresponding open-source 
  code is available.}. While extending the tree,
the configurations along the new edge are automatically projected onto $\mathcal{M}$. These projections 
are not very costly if the edge is close to the constrained manifold.

@@

\begin{algorithm}[h]
  \caption{Constrained-Extend($\mathcal{T},q_{near},q_{rand},f,\epsilon$)}
  \label{alg:constrained}
  \begin{algorithmic}
    \STATE {\color{red} // $I$ denotes the identity matrix.}
    \STATE $q_{rand}' \leftarrow q_{near} +   
    \left(I -  
    \left(\frac{\partial f}{\partial q}(q_{near})\right)^{+}
    \frac{\partial f}{\partial q}(q_{near})\right) (q_{rand} - q_{near})$
    \IF {$\texttt{constraint-solver}(q_{rand}',f,\epsilon) = q_{rand}''$} 
    \STATE RRT::Extend($\mathcal{T},q_{near},q_{rand}''$)
    \ENDIF
  \end{algorithmic}
\end{algorithm}

@@

This result has been long known and used in motion planning, in particular
for non-holonomic systems. A detailed proof can be found in
\cite{taix-94}. We will present a sketch of the proof to give an intuition about the 
corresponding algorithm.

@@

\subsubsection{Small-Time \textit{versus} Small-Space Controllability}
In the control theory literature, the property used is usually \textit{small-time controllability}, 
which states that for  all configurations  $q$, for  all times
$T>0$, the set of configurations accessible from $q$ in time less than
$T$ forms a  neighborhood of $q$. When accelerations and velocities are bounded,
small-time controllability implies small-space controllability. This is why 
a lot of motion planning previous work only refers to the sufficient small-time controllability
property. However, the reciprocate is not necessarily true:  a system can be 
small-space controllable and
not small-time, if the trajectories generated by its controller are arbitrarily long.
The important property, regarding
motion planning application, is small-space controllability, as 
Theorem~\ref{thm:ssc} shows. In the following, we show that legged robots
are small-space controllable, but  not that they are small-time controllable.
In fact, the  control method that we present does not follow the small-time controllability
property. For the sake of clarity, we have chosen to make the distinction between these
two controllability properties.

@@

This section  discusses walking robot small-space controllability. To clarify
the presentation, we consider a simplified model of a legged robot consisting of two feet
of zero mass and a point mass free to move in three dimensions.
We do not consider 
the kinematic chains between the feet and the mass. The robot is walking on a flat terrain,
and the feet are assumed to have a positive surface. For our presentation, it is not 
necessary to consider the feet height, so the configuration space of the robot is:
\[
\mathcal{C} = SE(2) \times SE(2) \times \mathbb{R}^3
\]
It is of dimension 9.
